Extreme heterogeneity of influenza virus infection in single cells

### 2018-01-11 paper/ResponseToReviews.tex

    b. Use simple models to illustrate how much variability can be accounted for by simple effects, such as expected Poisson co-infection frequency, or the likelihood of attaining full complement of RNP genes.

{\color{red}
Gini index is something we could definitely calculate. Not clear what we could compare it to, but we can at least calculate it and report it.
We then calculate Gini index under some simple model where everything is due to decreased transcription when missing segments.
Any such model is probably pretty bogus and oversimplified, so we'd want to do this in a way where everything ends up in a heavily caveated supplement that doesn't bog down text.
}

2. While the silent tagging method used to address co-infection is clever and appreciated, the issue is not fully settled. The dismissal of co-infection as a factor influencing cell-to-cell variation in lines 127-9 is based on too few cells to draw any conclusions, and thus needs to be tempered. Also, there is likely to be a significant amount of cryptic co-infection with identical barcode viruses (expected to be similar to that of mixed barcodes, ~10\%) that could influence measured heterogeneity. These points should be made in the text.

{\color{red}
The text will be revised to ensure that we don't say co-infection is unimportant. We're just saying that low MOI, it doesn't explain everything.

In addition, we will add the new Figure 4 - figure supplement 3 which shows HA or NS1 expression in cells infected at low or high MOI.
}

### 2018-01-11 paper/paper.tex
set up to add Figure 4 - Figure supplement 3, which shows HA / NS1 expression at high or low MOI

\begin{figure}[t!]
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_frac_flu_summary.pdf}}
\caption{\label{fig:viralbarcodes}
Synonymous barcodes on the viral mRNAs distinguish true infections from cells that contain viral mRNAs derived from leakage of lysed cells.
{\bf (A)}
Cells with at least two viral mRNAs for which the barcode could be called, arranged in order of increasing influenza transcript counts.
Bar heights denote the number viral mRNAs on a log\textsubscript{10} scale, bar coloring is linearly proportional to the fractions of viral mRNAs derived from wild-type and synonymously barcoded virus.
{\bf (B)}
Same as (A), but each bar is colored according to the relative fraction of the more common (major) and less common (minor) virus variant.
At low levels of viral mRNA there is often a roughly equal mix, suggesting contamination with viral mRNAs leaked from lysed cells.
At higher levels of viral mRNA, cells generally have only one viral variant, suggesting infection initiated by a single virion.
A few cells are also obviously co-infected with both viral variants.
{\bf (C)}
We determined a cutoff for calling ``true'' infections by finding the amount of viral mRNA per cell at which the viral barcode purity no longer increases with more viral mRNA.
The purity is the fraction of all viral mRNA in a cell derived from the most abundant viral barcode in that cell.
We fit a curve to the mean purity of all cells with more than the indicated amount of viral mRNA, and drew the cutoff at the point where this curve stopped increasing with the fraction of total mRNA derived from virus.
See the Methods for details.
{\bf (D)}
The number of cells identified as infected and co-infected for each sample, as well as the number of cells with any viral read.
For all subsequent analyses, we subsampled the number of uninfected cells per sample to the greater of 50 or the number of infected cells.
{\bf (E)} 
Distribution of the fraction of mRNA per cell derived from virus for both infected and co-infected cells.
}
\figsupp[Number of viral barcodes called.\label{figsupp:barcodescalled}]{
The number of viral barcodes called for each sample and gene segment. 
Viral transcripts are classified as \emph{syn} if they mapped to a synonymously barcoded influenza transcript, \emph{wt} if they mapped to a wild-type influenza transcript, \emph{invalid} if multiple reads for the same UMI differed on the status of the viral barcode, and as \emph{uncalled} if none of the reads for that UMI overlapped the region of the viral transcript containing the viral barcode.
For calculations of the number of reads in a cell derived from each viral barcode for each viral gene, the total number of detected molecules of that gene are multiplied by the fraction of those molecules with assignable barcodes that are assigned to that barcode.
}{\includegraphics[width=\linewidth]{figures/synbarcodes_umistats.jpg}}
\figsupp[Fraction of total viral mRNA derived from a given fraction of infected cells.\label{figsupp:cumulfracflu}]{
The total fraction of all viral mRNA among infected cells that is attributable to a given fraction of these cells.
For instance, the plot for the 8-hour sample shows that roughly 50\% of all viral mRNA is derived from 10\% of the infected cells.
}{\includegraphics[width=\linewidth]{figures/p_cumul_flu.pdf}}
\end{figure}

>>>>

\begin{figure}[t!]
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_frac_flu_summary.pdf}}
\caption{\label{fig:viralbarcodes}
Synonymous barcodes on the viral mRNAs distinguish true infections from cells that contain viral mRNAs derived from leakage of lysed cells.
{\bf (A)}
Cells with at least two viral mRNAs for which the barcode could be called, arranged in order of increasing influenza transcript counts.
Bar heights denote the number viral mRNAs on a log\textsubscript{10} scale, bar coloring is linearly proportional to the fractions of viral mRNAs derived from wild-type and synonymously barcoded virus.
{\bf (B)}
Same as (A), but each bar is colored according to the relative fraction of the more common (major) and less common (minor) virus variant.
At low levels of viral mRNA there is often a roughly equal mix, suggesting contamination with viral mRNAs leaked from lysed cells.
At higher levels of viral mRNA, cells generally have only one viral variant, suggesting infection initiated by a single virion.
A few cells are also obviously co-infected with both viral variants.
{\bf (C)}
We determined a cutoff for calling ``true'' infections by finding the amount of viral mRNA per cell at which the viral barcode purity no longer increases with more viral mRNA.
The purity is the fraction of all viral mRNA in a cell derived from the most abundant viral barcode in that cell.
We fit a curve to the mean purity of all cells with more than the indicated amount of viral mRNA, and drew the cutoff at the point where this curve stopped increasing with the fraction of total mRNA derived from virus.
See the Methods for details.
{\bf (D)}
The number of cells identified as infected and co-infected for each sample, as well as the number of cells with any viral read.
For all subsequent analyses, we subsampled the number of uninfected cells per sample to the greater of 50 or the number of infected cells.
{\bf (E)} 
Distribution of the fraction of mRNA per cell derived from virus for both infected and co-infected cells.
}
\figsupp[Number of viral barcodes called.\label{figsupp:barcodescalled}]{
The number of viral barcodes called for each sample and gene segment. 
Viral transcripts are classified as \emph{syn} if they mapped to a synonymously barcoded influenza transcript, \emph{wt} if they mapped to a wild-type influenza transcript, \emph{invalid} if multiple reads for the same UMI differed on the status of the viral barcode, and as \emph{uncalled} if none of the reads for that UMI overlapped the region of the viral transcript containing the viral barcode.
For calculations of the number of reads in a cell derived from each viral barcode for each viral gene, the total number of detected molecules of that gene are multiplied by the fraction of those molecules with assignable barcodes that are assigned to that barcode.
}{\includegraphics[width=\linewidth]{figures/synbarcodes_umistats.jpg}}
\figsupp[Fraction of total viral mRNA derived from a given fraction of infected cells.\label{figsupp:cumulfracflu}]{
The total fraction of all viral mRNA among infected cells that is attributable to a given fraction of these cells.
For instance, the plot for the 8-hour sample shows that roughly 50\% of all viral mRNA is derived from 10\% of the infected cells.
}{\includegraphics[width=\linewidth]{figures/p_cumul_flu.pdf}}
\figsupp[Flow cytometry analysis of viral protein production of cells infected at low and high infectious dose. \label{fig:flowdose}]
{
Expression of a viral protein in cells infected at high or low initial infectious dose (MOIs of 0.1 or 5 \jdbcomment{based on what titering method, TCID50?}).
While a higher dose leads to more cells expressing high amounts of viral protein, it does not substantially increase the amount of viral protein in either the low-expressing or high-expressing cells. 
Therefore, higher viral dose does not lead to a large continuous increase in viral protein production among all cells -- rather, it mostly changes the distribution of where cells fall in the highly heterogeneous distribution of viral protein production.
Cells dual-stained for NS1 and HA \jdbcomment{using what antibodies}.
The panels at left show the distribution of expression of NS1 in cells identified as virally infected based on the HA staining, and the panels at right show the distribution of HA expression in cells identified as virally infected based on the NS1 staining.
Data are shown for three independent replicates.
}{\jdbcomment{needs to be added}}
\end{figure}
